\section{Topology and challenges of the UK low-voltage power distribution network}
\label{ch-literature:sec:topology-of-lv-network}

Today's UK electricity network has grown over the past century and is based on an interconnected high-voltage gird.
The largest part of this grid is also known as the transmission network, which connects centralised power stations to distribution networks.
Those distribution networks supply electricity to all loads across the mainland of the UK, including industrial, urban and rural customers\footnote[1]{Some small and remote UK islands like the Shetland islands are not connected to this national grid and have their separate electricity infrastructure. Therefore they are not considered as part of this thesis since the study of this kind of network lies outside the research scope.}.

The entire structure of the network is a three-phase Alternating Current (AC) system since this allowed easy voltage level conversion with the use of transformers, i.e. without the need of power electronics.
In the UK, the highest voltage level for generation and transmission is 400kV.
Such a high voltage requires a relatively small current to transmit the generated bulk power, which in turn reduces conduction losses and maximises the efficiency of the high-voltage network.
Regional supply points step-down this high voltage to 132kV\footnote[1]{In some cases regional supply points provide 127kV instead of 132kV.} to deliver power to Distribution Network Operators (DNOs).
From the primary level of the distribution network and onwards, this so called medium-voltage is stepped down to 33kV, then 11kV and finally 400V, in order to cater for heavy industry, medium clients and household sized customers, respectively.

\nomenclature[G]{DER}{Distributed Energy Resource}
\nomenclature[G]{ADMD}{After Diversity Maximum Demand}
\nomenclature[G]{OLTC}{On-Line Tap-Changer}

Primary substations in the UK are equipped with regulation equipment, e.g. On-Line Tap-Changers (OLTC), to increase or decrease the voltage on the secondary transformer side depending on the current level of demand.
Secondary transformers do not have such regulating equipment and instead apply a constant voltage conversion ratio which is set according to the network's typical demand.
This aim of this network regulation is to keep distribution level voltages within their statutory operating bands, i.e. 230V +10\% -6\% for the LV network as specified by the Electricity Supply Quality and Continuity Regulation (ESQRC) \cite{HealthandSafetyExecutive2002} and Engineering Recommendation G59 \cite{EnergyNetworksAssociation2013}.
In the UK, all households are connected to one of the three phases of the 230V distribution network.
To achieve a balanced network, each customer's phase allocation is chosen at random.
Also, throughout the majority of the power network's development period, customers were light consumptive loads, which meant that a reasonably predictable power flew from higher voltage levels towards the lower voltage levels.
Therefore, traditional network planning approaches to circumvent constraint violations, follow the commonly used practice of aggregating a large number of customers and designing the power delivery network to cater for their largest probable demand, i.e. the After Diversity Maximum Demand (ADMD) method \cite{Richardson2010a}.
However, this ADMD method has remained the same for many years and uses historical load analysis and standard growth assumptions, which are both no longer valid in this unprecedented LCT uptake scenario \cite{Yunusov2016}.

Firstly, because the injection of power from Distributed Energy Resources (DERs), e.g. rooftop solar PV, can reverse the power flow, and secondly, because large and volatile electrified products, e.g. home-charging Electric Vehicles (EVs), are predicted to significantly increase demand at peak times.
Such technologies have significant impact on the voltage stability \cite{Petinrin2016}, and combined with the expected increased in phase unbalance, traditional network management methods may no longer be able to effectively mitigate their negative impact.
This means that in situ equipment needs to be managed actively via innovation in the use of existing and new technologies; otherwise not only the frequency of constraint violations will increase, but also the frequency of service disruptions and customer minutes lost is expected to rise alongside the proliferation of LCTs \cite{Ault2008a}.
One such innovative technology, which is the main focus of the presented research, is the installation and management of battery storage \cite{Chen2009}, which is reviewed in the following section.















