%\setlength{\topmargin}{-.5in}

\epigraph{\textit{It is not a dream. It is a simple feat of scientific electrical engineering. Electric power can drive the world's machinery without the need of coal, oil or gas. Although perhaps humanity is not yet sufficiently advanced to be willingly lead by the inventor's keen searching sense. Perhaps it is better in this present world of ours where a revolutionary idea may be hampered in its adolescence. All this that was great in the past was ridiculed, condemned, combatted, suppressed only to emerge all the more triumphantly from the struggle. [...] Our duty is to lay the foundation for those who are to come and to point the way, yes humanity will advance with giant strides. We are whirling through endless space with an inconceivable speed, all around everything is spinning, everything is moving, everywhere there is energy.}}{--- Nicola Tesla}


\chapter*{Abstract}

\addcontentsline{toc}{chapter}{Abstract}


%enter text for the abstract below

%% version 3
British Distribution Network Operators (DNOs) are facing challenges due to the\linebreak
energy sectors transitioning into a low carbon economy. This thesis aims to present\linebreak
novel methods to aid DNOs in operating their Low-Voltage (LV) networks despite this\linebreak
ongoing transition and its entailed challenges. The presented methods are realised\linebreak
with the use of Battery Energy Storage Solutions (BESS) and they develop BESS\linebreak
energy management algorithms whilst focusing on communication regimes and\linebreak
sub-half-hourly volatility in demand. Consequently, improving LV network operation\linebreak
mainly considers the reduction of peak power flow, but also includes reducing\linebreak
energy losses, voltage deviation, the magnitude of neutral currents and phase un-\linebreak
balance. Without these methods, DNOs would have to rely on traditional network re-\linebreak
enforcements so that LV networks are kept within e.g. statutory voltage bands.\linebreak
Extending current literature by developing these methods to control a single energy\linebreak
resource and a distributed BESS - whilst considering requirements for communication\linebreak
systems that may effect BESS control - is the main contribution of this thesis.

%% version 2
%British national low-carbon targets have resulted in a transition within the UK\linebreak
%energy and transport sectors. This transition entails the development and uptake of\linebreak
%new Low-Carbon Technologies (LCTs). Some technologies (e.g. photovoltaic) can\linebreak
%offer demand alleviating services since they can supply electricity to the grid,\linebreak
%whilst others (e.g. Electric Vehicles - EVs) will increase demand for grid supplied\linebreak
%electricity. Distribution System Operators (DNOs), having to keep the distribution\linebreak
%system within statutory constraints, are thus facing significant challenges like\linebreak
%more volatility in demand, higher voltage deviation and larger phase unbalance.\linebreak
%Energy storage devices can be a feasible alternative to conventional reinforcement\linebreak
%activities, since they can be managed actively. Developing novel control for\linebreak
%Battery Energy Storage Solutions (BESS) is thus the main focus of this thesis.\linebreak
%Extending current literature by developing control algorithms for a single \linebreak
%energy system and distributed BESS whilst taking into account the communication\linebreak
%requirements for effective control is its main contribution.

%% version 1
%Beginning the transition of UK energy systems to reach national low-carbon\linebreak
%economy targets is expected to put significant strain onto the existing power\linebreak
%networks. Reinforcing the Low-Voltage (LV) distribution network, to assure its\linebreak
%operation remains within statutory constraints, will become essential and necessary\linebreak
%when e.g. Low-Carbon Technologies (LCTs) like heat pumps, photovoltaic panels or\linebreak
%Electric Vehicles (EVs) begin penetrating the network. The expected impact\linebreak
%such as higher voltage deviation, phase unbalance, neutral current and larger\linebreak
%mean power flow, will make the Distribution Network Operator's (DNO's) role of\linebreak
%ensuring their LV networks remain within the aforementioned statutory limits \linebreak
%a significant challenge. The deployment of Battery Energy Storage Solution \linebreak
%(BESS) is seen as a feasible alternative to conventional network reinforcement.\linebreak
%Energy storage control, for both single, DNO owned and multiple, privately owned\linebreak
%energy storage devices, to improve LV network operation without the need for any\linebreak
%reinforcement, is the main contribution of this thesis.

The first BESS control algorithms that are developed throughout this thesis are designed to incorporate both half-hourly forecasts and sub-half-hourly load volatility in order to take into account both the long term demand trends and its high volatility.
As a result, key network parameters and their interplay could be identified and daily load peaks that are caused by load volatility could on average be reduced by an additional 3.8kW, even when the underlying demand forecast was erroneous.
These methods are developed and address challenges for controlling a single (i.e. DNO owned) BESS - hence neglecting the challenges that are based on inter-device communication.
Those neglected challenges are addressed in the subsequent BESS control methods where a desynchronised Multi-Agent Network (MAS) and communication less BESS control fill this gap.
Results show how internal algorithm behaviour changes when desynchronising the communication environment, but without impacting the global performance of the distributed BESS.
Also, real-time performance of the communication less control algorithm is studied on a time-series and statistical basis and shows how effects from uncoordinated Low-Carbon Technologies (LTCs) i.e. Electric Vehicle (EV) charging, can be successfully mitigated.
All key objectives that align with the aforementioned achievements have been met for each of the presented contribution and the comparable storage control techniques in literature are either met or exceeded in performance when subjected to the available datasets.
 