\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to begin by thanking my two PhD supervisors and tutors Dr Ben Potter and Professor William Holderbaum for their continuing support, help and guidance throughout my PhD.
In addition, I would like to thank the \textit{University of Reading} and sponsors of the \textit{NTVV Project} for providing the funding for my research.
Also, I would like to thank the \textit{ITNG} members of the former \textit{School of Systems Engineering} for providing and maintaining the vital computational resources, and I would like to thank our industrial project partner, \textit{Scottish and Southern Electricity Networks}, for providing thei network models and datasets; both were were vital necessities to conduct and complete my research.

I would like to thank members of the \textit{School of the Built Environment}, the \textit{School of Biological Sciences} and the \textit{School of Psychology and Clinical Language Sciences} for taking their time to share, exchange and discuss ideas and for challenging my thinking and expertise throughout the journey towards the PhD.
In particular I would like to express special thanks to individuals in the \textit{Energy Research Laboratory}: Dr Timur Yunusov, Ana Rodriguez-Arguelles and Peter Adams; individuals from the \textit{SUSPORTS Project}: Dr Ian Harrison and Stefano Pietrosanti; an individual in the \textit{Brain Embodiment Laboratory}: Asad Malik; and individuals at the \textit{TSBE Centre} and the \textit{School of Construction Management and Engineering}: Alice Gunn and Vicky Papaioannou.

There are many current and past colleagues, researchers and friends who have not only technically aided my research, but have also reminded me of life outside academia.
Through their willingness to listen, motivate, support and (sometimes) distract me, they have made my time as a PhD student considerably more enjoyable; so thank you.

%To name a few: Thomas Chung, Alan Halford, Nicholas Holley, Joshua Eadie, Mary Pothitou, Saadia Ansari, Dr Laura Daniels, Deborah Ritzmann, Dr Dan Williams, Dr Stephen Haben, Dr Christopher Shackleton and Dr Charles Moorey.

Last but not least, I would like to express my greatest gratitude and thanks to my family for all of their support and understanding throughout my PhD and time at university; it is much appreciated and I could not have done it without you.