\section{Results and Discussion}
\label{ch3:sec:results}

\subsection{Algorithm performance for synchronised operation}
\label{ch3:subsec:algorithm-performance-synchronised}

The objective of the smart-charging algorithm is to distribute the charging demand of a fleet of EVs over the underlying base\hladd{line load}\hlrem{ demand} in such a way that no additional demand spikes are produced.
After assigning each EV's energy demand to its initially known demand trough, the algorithm produces a new demand spike since all EVs are charging simultaneously.
Through repetitive iterations and reallocating a portion of the assigned energy to different demand troughs, the algorithm is then able to spread all EVs' demands to form a flat demand profile in the end.
This process is shown in Figure~\ref{ch3:fig:time-series}.

\input{_chapter3/fig/time-series/time-series}

\input{_chapter3/fig/oscillation/oscillation}

Here, the first algorithm iteration is shown in Figure~\ref{ch3:subfig:time-series-1}, where allocated power profile produces two new morning spikes of around 200kW and subsequently 110kW.
The second iteration however reduces these spikes by the factor $\alpha$ (i.e $0.2$) and redistributes the undone charging powers over the new power profile.
Figure~\ref{ch3:subfig:time-series-2} shows this reduction and reallocation.
Figure~\ref{ch3:subfig:time-series-3} is the third iteration that reduces and redistributes the peaks even further.
In the end (i.e. when $n=100$) the resulting power profile becomes as flat as possible which is shown in Figure~\ref{ch3:subfig:time-series-last}.
Throughout these iterations, it can be observed how the peak load in the total power (i.e. $\textbf{p}_n$) reduces and it can be observed how the changes in charging power (i.e. $\textbf{p}_{n+1}-\textbf{p}_n$) reduce in variance, which indicates that the algorithm works for the chosen parameters of $\alpha$ and $\beta$.
However, different parameters of $\alpha$ and $\beta$ do impact the performance of this synchronised algorithm execution as shown in Figure~\ref{ch3:fig:oscillation}.

\input{_chapter3/tab/pair-comparison}

Whereas the $\alpha$ and $\beta$ parameters use to produce the results in Figure~\ref{ch3:fig:time-series} reduced the power spike, those parameters in \ref{ch3:fig:oscillation} where $\alpha = \beta = 1.0$ did not.
In fact, an oscillating behaviour can be observed since the initially applied power profile is completely undone and completely reassigned onto a different demand trough.
Since this produces similar peaks, the same procedure repeats and reassigns the complete power profile back to the original demand troughs.
In the end, these charging spikes can never be fully mitigated and the algorithm did not smoothen the total demand.
The resulting charging spike of more than 200kW could therefore cause significant issues if \hlrem{e.g. }the underlying physical network has not been scaled appropriately\hladd{, for example}.
This longevity of issue becomes more evident when comparing the $\zeta^\text{PAR}$ and $\zeta^\text{TRA}$ values for both parameter pairs.
The evolution of $\zeta^\text{PAR}$ and $\zeta^\text{TRA}$, as tabulated in Table~\ref{ch3:tab:pair-comparison}, shows this difference in performance and convergence of the algorithm when subjected to different values of $\alpha$ and $\beta$.
Those figures indicate that a well chosen pair of $\alpha$ and $\beta$ results in a convergence value (i.e. $b$) that is greater than zero.
No convergence occurred however for those values of $\alpha$ and $\beta$ that resulted in the oscillating behaviour (i.e. convergence rate $b$ is zero).

Next, the entire range of $\alpha$ and $\beta$ needed to be studied since different parameter pairs are likely to result in different convergence rates (and therefore different algorithm performance).
These results (i.e. for the synchronised algorithm performance) are plotted in Figure~\ref{ch3:fig:all-sync}.

\input{_chapter3/fig/all-sync/all-sync}

Figure~\ref{ch3:subfig:all-sync-par} and Figure~\ref{ch3:subfig:all-sync-tra} show how the final values for both $\zeta^\text{PAR}$ and $\zeta^\text{TRA}$ were lowest when either $\alpha$ or $\beta$ was chosen closer to zero.
This result coincides with the finding that hard reduction and reallocation lead to an oscillating behaviour of the algorithm.
Similarly, the convergence of those two performance metrics, as shown in Figure~\ref{ch3:subfig:all-sync-par-conv} and Figure~\ref{ch3:subfig:all-sync-tra-conv}, was highest when $\alpha$ approached one and $\beta$ approached zero.
This behaviour is by design since a larger value of $\alpha$ increases the rate at which the currently applied peak is reduced whilst a smaller value of $\beta$ limits the amount that can be reallocated for each time slot.
Such clear behavioural differences for different pairs of $\alpha$ and $\beta$ indicate an optimal operation region of the algorithm within the \hlrem{north-eastern}\hladd{top right} quadrant of the plot.
\hladd{This quadrant indicates optimal operation since low $\zeta_\text{PAR}$ and $\zeta_\text{TRA}$ values are achieved (see Fig.}~\ref{ch3:subfig:all-sync-par}\hladd{ and Fig.}~\ref{ch3:subfig:all-sync-tra}\hladd{) at a relatively high rate of convergence (see Fig.}~\ref{ch3:subfig:all-sync-par-conv}\hladd{ and Fig.}~\ref{ch3:subfig:all-sync-tra-conv}\hladd{).}
Whether the algorithm still performs in this way when introducing desynchronisation is answered in the subsequent section, Section~\ref{ch3:subsec:algorithm-performance-desynchronised-regular}.


\subsection{Algorithm performance for desynchronised operation with regular timing}
\label{ch3:subsec:algorithm-performance-desynchronised-regular}

\input{_chapter3/fig/time-series-desync/time-series-desync}

Looking at the evolution of the time-series when desynchronising the algorithm's execution shows significant differences already.
Figure~\ref{ch3:fig:time-series-desync} shows thie evolution for the same parameters as those chosen for Figure~\ref{ch3:fig:time-series}.
The difference is however that the assignment of charging powers lead to a significantly lower demand spike at the very beginning of executing the algorithm.
Subsequent iterations then reduce this spike much broader than it has been the case when executing the algorithm in a synchronised manner.
Therefore, more demand troughs are filled and a smoother profile is obtained\hladd{ when compared to sub-optimal pairs of $\alpha$ and $\beta$.
It should however be noted that the valley filling is not as effective, but} much quicker\hladd{,} when compared to the synchronised case.
\hladd{After all, the initial power spike already reduced to below 100kW, whilst the synchronised cases always started at peaks above 200kW.}

\input{_chapter3/fig/all-async-regular/all-async-regular}

This behaviour becomes particularly apparent when looking at the full range of $\alpha$ and $\beta$ values.
Figure~\ref{ch3:fig:all-async-regular} shows the same full range analysis as the figure in the previous section did, i.e. Figure~\ref{ch3:fig:all-sync}.
When comparing them at the same scale, $\zeta^\text{PAR}$ and $\zeta^\text{TRA}$ values (plotted in Figure~\ref{ch3:subfig:all-async-regular-par} and Figure~\ref{ch3:subfig:all-async-regular-tra}, respectively) have significantly lowered in magnitude.
This indicates a much better performance of the algorithm across the entire range of $\alpha$ and $\beta$ parameters.
Convergence rates (i.e. plotted in Figure~\ref{ch3:subfig:all-async-regular-par-conv} and Figure~\ref{ch3:subfig:all-async-regular-tra-conv}) were however not impacted to the same extend.
This indicates that the underlying execution of the algorithm still performs as intended, but the interplay between the agents that implement this algorithm changes the outcome of the aggregated result.

The next step is to assess whether desynchronising the algorithm's execution by randomising the loop delays yields any further changes in algorithm performance and behaviour.
Results from that step are presented in the following section, Section~\ref{ch3:subsec:algorithm-performance-desynchronised-irregular}.

\subsection{Algorithm performance for desynchronised operation with irregular timing}
\label{ch3:subsec:algorithm-performance-desynchronised-irregular}

\input{_chapter3/fig/time-series-desync-irregular/time-series-desync-irregular}

As shown in Figure~\ref{ch3:fig:time-series-desync-irregular}, the difference between regular and irregular loop delays when executing the smart-charging algorithm is barely noticeable.
The interlaced querying still causes each agent to react to a slightly different network demand profile which results in a varied power profile allocation.
A functioning peak reduction behaviour is therefore a positive sign since this irregular algorithm desynchronisation (as introduced in Section~\ref{ch3:subsec:desynchronisation}) represents the worst algorithm deployment scenario.
Performance and convergence do however need to be inspected for the complete range of $\alpha$ and $\beta$ values so that the results can be compared to the previous findings.

\input{_chapter3/fig/all-async-irregular/all-async-irregular}

Figure~\ref{ch3:fig:all-async-irregular} shows the results for this range of $\alpha$ and $\beta$ values when executing the algorithm on a desynchronised MAS with irregular loop delays.
The values for $\zeta^\text{PAR}$ and $\zeta^\text{TRA}$ are still significantly lower than they were for the synchronised case, but they do not differ much from the regular desynchronisation case.
The same is true when comparing convergence which indicates that the algorithm's underlying execution still performs as intended.
