\chapter{Introduction}
\label{ch-introduction}

The aim of the work that is presented in this thesis is to make a contribution in control of Battery Energy Storage Systems (BESS) that can aid Distribution Network Operators (DNOs) in improving the operation and reliance of their Low-Voltage (LV) networks.
By adjusting sub-half-hourly operation and communication regimes of the BESS grid operation, performance is improved by mainly considering peak power flow - but performance criteria also include \hlrem{e.g. }voltage deviation, phase imbalance, distribution losses and the magnitude of neutral currents.

Providing BESS based network support is expected to become a vital necessity since the predicted increase in electricity demand and its volatility will negatively effect the performance of the UK distribution networks.
Due to the network's design and topology higher and more volatile demand is predicted to cause issues including voltage deviation, asset overloads, equipment damage and (in the worst case) service disruptions.
As discussed in this thesis, BESS is a suitable alternative to traditional network reinforcements.
However, successfully combining fast system response capabilities (i.e. at sub-half-hourly resolution) with traditional operation schedules (i.e. at half-hourly resolution) to yield the best impact on network performance (\hlrem{e.g.}\hladd{for example} peak-reduction) is still an open research challenge.
Also, with the proliferation of household-connected BESS and Electric Vehicles (EVs), the development of control methods (i.e. in respect to the communication and synchronisation requirements that are needed to coordinate BESS operation) is also an ongoing research challenge.

In this chapter, Chapter~\ref{ch-introduction}, the background and motivation for the conducted research are presented.
Then, on the basis of the identified challenges and opportunities for battery energy storage in the electricity distribution network, the problem statement and all research objectives are outlined.
At the end of this chapter, all contributions and publications are summarised, and the structure of the rest of this thesis is presented.

\input{_introduction/sec/background}

\input{_introduction/sec/problem-statement}

\input{_introduction/sec/contributions}

\input{_introduction/sec/publications}

\input{_introduction/sec/thesis-structure}